Light and growth form interact to shape stomatal ratio among British angiosperms

### 2017-09-26 ms/ms.Rnw
updated ms
adding phylo figure

I obtained data on ab- and adaxial stomatal density on \Sexpr{nrow(stomata_unfiltered)} species from British Ecological Flora \citep{Salisbury_1927, Fitter_Peat_1994a, BEF}. Following recent comparative analyses \citep[e.g.][]{Bartelheimer_Poschlod_2016, Salguero-Gomez_etal_2016, Shipley_etal_2017}, I used Ellenberg light indicator values \citep{Ellenberg_1974} and Raunki\ae r life form \citep{Raunkiaer_1934} as measures of light habitat and growth form, respectively. Hence, I am assuming that the species' light habitat is closely related to the type of habitat (open versus closed) where that species is found. Both attributes have been recently updated by taxonomic experts of the British flora (PLANTATT, \cite{Hill_etal_2004}). Ellenberg light indicator values are hereafter abbreviated \el. I used a dated molecular phylogeny of the British flora \citep{Lim_etal_2014} available from TreeBASE (http://treebase.org/; accession number 15105). \Sexpr{nNotInPhy1} species (\Sexpr{round(100 * nNotInPhy1 / nrow(stomata_unfiltered), 1)}\%) in the dataset were not present in the phylogeny. For \Sexpr{nNotInPhy1 - nNotInPhy2} of these species, I used the position a congeneric species as a proxy for the focal species \citep[following][]{Pennell_etal_2016}. When multiple congeneric species were present, I consulted the phylogenetic literature to identify the most closely related proxy species \citep{Scheen_etal_2004, Salmaki_etal_2013}. For the remaining \Sexpr{nNotInPhy2} missing species, I positioned them in the tree based on phylogenetic relationships to other genera or families present in the tree \citep{Fior_etal_2006}. Because many phylogenetic comparative methods do not allow polytomies, zero-length branches, and non-ultrametric trees, I made several small adjustments to the tree. I resolved polytomies randomly using the `multi2di' function in \pkg{phytools} version 0.6-00 \citep{Revell_2012}. I added 0.02 my to all zero-length branches, as this was approximately the length of the shortest nonzero branch length in the tree. After these changes, I slightly altered terminal branch lengths to make the tree precisely ultrametric.

@@

I tested whether Raunki\ae r life form was associated \el~among British angiosperms using ANOVA with Type-2 sum of squares. I did not use phylogenetic ANOVA for this test because there was no phylogenetic signal in the regression fit using \pkg{phylolm} version 2.5 \citep{Ho_Ane_2014}. See the R code accompanying this paper for further detail. I predicted that species with faster life histories, especially therophytes (annuals), would have greater \els~than species with slower life histories, especially phanerophytes, which are mostly long-lived trees.

>>>>

I obtained data on ab- and adaxial stomatal density on \Sexpr{nrow(stomata_unfiltered)} species from British Ecological Flora \citep{Salisbury_1927, Fitter_Peat_1994a, BEF}. Following recent comparative analyses \citep[e.g.][]{Bartelheimer_Poschlod_2016, Salguero-Gomez_etal_2016, Shipley_etal_2017}, I used Ellenberg light indicator values \citep{Ellenberg_1974} and Raunki\ae r life form \citep{Raunkiaer_1934} as measures of light habitat and growth form, respectively. Hence, I am assuming that the species' light habitat is closely related to the type of habitat (open versus closed) where that species is found. Both attributes have been recently updated by taxonomic experts of the British flora (PLANTATT, \cite{Hill_etal_2004}). Ellenberg light indicator values are hereafter abbreviated \el. I used a dated molecular phylogeny of the British flora \citep{Lim_etal_2014} available from TreeBASE (http://treebase.org/; accession number 15105). \Sexpr{nNotInPhy1} species (\Sexpr{round(100 * nNotInPhy1 / nrow(stomata_unfiltered), 1)}\%) in the dataset were not present in the phylogeny. For \Sexpr{nNotInPhy1 - nNotInPhy2} of these species, I used the position of a congeneric species as a proxy for the focal species \citep[following][]{Pennell_etal_2016}. When multiple congeneric species were present, I consulted the phylogenetic literature to identify the most closely related proxy species \citep{Scheen_etal_2004, Salmaki_etal_2013}. For the remaining \Sexpr{nNotInPhy2} missing species, I positioned them in the tree based on phylogenetic relationships to other genera or families present in the tree \citep{Fior_etal_2006}. Because many phylogenetic comparative methods do not allow polytomies, zero-length branches, and non-ultrametric trees, I made several small adjustments to the tree. I resolved polytomies randomly using the `multi2di' function in \pkg{phytools} version 0.6-00 \citep{Revell_2012}. I added 0.02 my to all zero-length branches, as this was approximately the length of the shortest nonzero branch length in the tree. After these changes, I slightly altered terminal branch lengths to make the tree precisely ultrametric.

@@

I tested whether Raunki\ae r life form was associated with \el~among British angiosperms using ANOVA with Type-2 sum of squares. I did not use phylogenetic ANOVA for this test because there was no phylogenetic signal in the regression fit using \pkg{phylolm} version 2.5 \citep{Ho_Ane_2014}. See the R code accompanying this paper for further detail. I predicted that species with faster life histories, especially therophytes (annuals), would have greater \els~than species with slower life histories, especially phanerophytes, which are mostly long-lived trees.
